---
title: "p8105_hw_fwt2107"
author: "Felix Tran"
date: "September 16, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
## Creating the dataframe
We will create a dataframe with 4 entries for problem 1. We initialize the 4 entries first.

random_sample is a vector of 10 random numbers from a uniform distribution [0, 5].

logic_vector is a vector showing which numbers in random_sample are greater than 2.

char_vector is a character vector of length 10.

factor_vector is a factor of length 10.

```{r}
random_sample <- runif(10, min = 0, max = 5)
logic_vector <- 2 > random_sample
char_vector <- vector(mode = "character", length = 10)
factor_vector <- factor(c(vector(length = 10)))
```


prob1_dataframe is created containing the 4 vectors

```{r}
library(tibble)
prob1_dataframe <- data_frame(
  random_sample,
  logic_vector,
  char_vector,
  factor_vector
)
```

## Calculating the means of each vector
The mean can be calculate for random_sample and logic_vector because their
values are either numeric or logical 

```{r}
mean(prob1_dataframe$random_sample)
mean(prob1_dataframe$logic_vector)
```


However, the mean cannot be calculated for char_vector and factor_vector 
because their values are not numeric nor logical
```{r}
mean(prob1_dataframe$char_vector)
mean(prob1_dataframe$factor_vector)
```

## Converting variable types
Applying the as.numeric function to logic_vector converts the TRUE values
into 1's and the FALSE values into 0's. Applying the as.numeric function to
char_vector converts all the "" entries into NA's. Applying the as.numeric
function to factor_vector converts all the values into 1's.
```{r, results = "hide"}
as.numeric(logic_vector)
as.numeric(char_vector)
as.numeric(factor_vector)
```

We convert char_vector from character vector to factor to numeric. This
gives us a numeric vector with all entries = 1. We also convert
factor_vector from factor to character vector to numeric. This gives
us a vector with all entires NA.
```{r}
as.numeric(as.factor(char_vector))
as.numeric(as.character(factor_vector))
```


# Problem 2
## Creating the dataframe

We first create the vectors we will eventually combine for our dataframe.

x is a random sample of 1000 numbers from the standard normal distribution.
y is a random sample of 1000 numbers from the standard normal distribution.
logic_vector is a vector of logic values indicating if x + y > 0.
num_coerced is logic_vector coereced as a numeric vector.
factor_coerced is logic_vector coerced as a factor.
```{r}
x <- rnorm(1000)
y <- rnorm(1000)
logic_vector <- (x + y) > 0
num_coerced <- as.numeric(logic_vector)
factor_coerced <- as.factor(logic_vector)
```

We create the dataframe containing the vectors we just created.
```{r}
plot_df <- data_frame(
  x,
  y,
  logic_vector,
  num_coerced,
  factor_coerced
)
```

This dataframe is named plot_df. It has `r nrow(plot_df)` rows and 
`r ncol(plot_df)` columns. The mean of x is `r mean(x)`, and the median
of x is `r median(x)`. The proportion of cases in logic_vector which are
TRUE is `r (length(which(logic_vector == T)))/10`%. 

This is the first of 3 scatterplots of y vs x. The points with coordinates
(x, y) in which x + y > 0 are one color, and the remaining points are a
different color. These 2 colors are the only 2 colors used in the color 
scale.
```{r}
library('ggplot2')
logic_ggplot <- ggplot(plot_df, aes(x = x, y = y, color = logic_vector)) + 
  geom_point()
logic_ggplot
```

This is the second scatterplot of y vs x. The points with coordinates
(x, y) in which x + y > 0 are one color, and the remaining points are a
different color. The color scale uses a gradient, but because there are
only two different numeric values in num_coerced only two shades of the
scale are used.
```{r}
library('ggplot2')
num_ggplot <- ggplot(plot_df, aes(x = x, y = y, color = num_coerced)) + 
  geom_point()
num_ggplot
```

This is the third scatterplot of y vs x. The points with coordinates
(x, y) in which x + y > 0 are one color, and the remaining points are a
different color. Like the first scatterplot which used logic_vector for 
its color scale, these 2 colors are the only 2 colors used in the color 
scale.
```{r}
library('ggplot2')
factor_ggplot <- ggplot(plot_df, aes(x = x, y = y, color = factor_coerced)) + 
  geom_point()
factor_ggplot
```