---
title: "p8105_hw_fwt2107"
author: "Felix Tran"
date: "September 16, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
## Creating the dataframe
We will create a dataframe with 4 entries for problem 1. We initialize the 4 entries first.

random_sample is a vector of 10 random numbers from a uniform distribution [0, 5].

logic_vector is a vector showing which numbers in random_sample are greater than 2.

char_vector is a character vector of length 10.

factor_vector is a factor of length 10.

```{r}
random_sample <- runif(10, min = 0, max = 5)
logic_vector <- 2 > random_sample
char_vector <- vector(mode = "character", length = 10)
factor_vector <- factor(c(vector(length = 10)))
```


prob1_dataframe is created containing the 4 vectors

```{r}
library(tibble)
prob1_dataframe <- data_frame(
  random_sample,
  logic_vector,
  char_vector,
  factor_vector
)
```

## Calculating the means of each vector
The mean can be calculate for random_sample and logic_vector because their
values are either numeric or logical 

```{r}
mean(prob1_dataframe$random_sample)
mean(prob1_dataframe$logic_vector)
```


However, the mean cannot be calculated for char_vector and factor_vector 
because their values are not numeric nor logical
```{r}
mean(prob1_dataframe$char_vector)
mean(prob1_dataframe$factor_vector)
```

## Converting variable types
Applying the as.numeric function to logic_vector converts the TRUE values
into 1's and the FALSE values into 0's. Applying the as.numeric function to
char_vector converts all the "" entries into NA's. Applying the as.numeric
function to factor_vector converts all the values into 1's.
```{r, results = "hide"}
as.numeric(logic_vector)
as.numeric(char_vector)
as.numeric(factor_vector)
```

We convert char_vector from character vector to factor to numeric. This
gives us a numeric vector with all entries = 1. We also convert
factor_vector from factor to character vector to numeric. This gives
us a vector with all entires NA.
```{r}
as.numeric(as.factor(char_vector))
as.numeric(as.character(factor_vector))
```


